---
title: "cluster_analytics"

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
  encoding=encoding,
  output_file=file.path(dirname(input_file), out_dir, 'index.html'))})

author: "Catherina Cader"
date: "11 4 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Overview 

This document provides an overview of the population cluster derived from the High Resolution Layer (<https://www.ciesin.columbia.edu/data/hrsl/>) and OSM (<https://www.openstreetmap.org/data>) for the case of Nigeria.
The data processing and analysis is done using R. This document provides a transparent documentation on the processes and results.

For the analysis the follwing packages are required:
```{r echo=FALSE}
#required packages

library(raster)
library(sf)
library(sp)
library(rgdal)
library(rgeos)
library(dbscan)
library(dplyr)
library(ggplot2)
library(mapview)
library(tmap)
library(geosphere)
library(leaflet)

```


## Input data 

Three data set are used:
*HRSL
*OSM
*Admin boundaries

```{r}

#read in spatial data

datapath<-("\\\\srv02\\RL-Institut\\04_Projekte\\240_NESP2\\03-Projektinhalte\\02_Data\\02_Data_from_online_sources\\05_Population_data\\hrsl_nga_v1")
builtup<-raster(file.path(datapath,"hrsl_nga_settlement.tif"))
adm1<-st_read("\\\\srv02\\RL-Institut\\04_Projekte\\240_NESP2\\03-Projektinhalte\\02_Data\\02_Data_from_online_sources\\04_Administrative_Boundaries\\nga_admbnda_adm1_osgof")

adm1 %>% select(admin1Name) %>% ggplot() + geom_sf() + labs(title = "Nigeria")

adm1 %>% filter(admin1Name == "Kano") %>% ggplot() + geom_sf() + labs(title = "Kano")

state_list<-list()


```


## Data processing

Extract the population raster data for each state

```{r echo=FALSE}
Kano<- adm1[adm1$admin1Name=='Kano', ]

builtup_Kano<-crop(builtup,Kano) #crop clips the data to the given extent
builtup_Kano<-mask(builtup_Kano,Kano) #mask excludes data inside the extent but outside the specified shape

#transform raster centerpoints to points
pts_buildup_Kano<-rasterToPoints(builtup_Kano, spatial=TRUE)

summary_Kano <-summary(pts_buildup_Kano)

input=as.data.frame.array(pts_buildup_Kano@coords)


```

## Clustering

The DBSCAN algorithm is used to cluster the points based on the distance to each other and a minimum number of points per cluster.

```{r}

test=dbscan(input, eps = 0.00083, minPts = 3, weights = NULL)
test
input$cluster <- test$cluster

#split the original data into two according to whether dbscan has assigned or cluster or noise.

groups  <- input %>% filter(cluster != 0)
noise  <- input %>% filter(cluster== 0)

out <- input %>% 
  st_as_sf(coords = c("x","y"), crs = 4326) %>%
  group_by(cluster) %>% 
  summarise() %>% 
  st_convex_hull()

out %>% filter(cluster == 0) 

outfiltered <- out %>% filter(cluster != 0)
union <- st_union(outfiltered)



```

## Visualization

The resulting clusters are shown in the map below.


```{r}

mapview(st_geometry(powerlines$geometry))+
mapview(union[[1]], color= "red", col.regions = "red")
  

# Transform the sfc object to polygons

poly <- st_collection_extract(union, type = "POLYGON", warn = FALSE)
mapview(poly, color= "red", col.regions = "red")+
  mapview(st_geometry(powerlines$geometry))




#Calculate centroids of the polygons:

poly$centroids <-
  st_centroid( x = poly ) %>%
  st_geometry()

```
## Distance calculations

Between the polygons and grid infrastructure


```{r}

powerlines<-st_read("\\\\srv02\\RL-Institut\\04_Projekte\\240_NESP2\\03-Projektinhalte\\02_Data\\02_Data_from_online_sources\\06_power_infrastructure\\columbia_grid_tracking\\kedco-grid-data-2016-shapefiles\\KEDCO_Grid_Data_Simplified_06172015_with_overlaps_removed.shp")

# Calculate distances -----------------------------------------------------

dist <- st_distance(centroids, st_cast(st_geometry(powerlines), "POINT"), by_element = TRUE) 

```