---
title: "cluster_analytics"

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
  encoding=encoding,
  output_file=file.path(dirname(input_file), out_dir, 'index.html'))})

author: "Catherina Cader"
date: "11 4 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Overview 

This document provides an overview of the population cluster derived from the High Resolution Layer (<https://www.ciesin.columbia.edu/data/hrsl/>) and OSM (<https://www.openstreetmap.org/data>) for the case of Nigeria.
The data processing and analysis is done using R. This document provides a transparent documentation on the processes and results.

For the analysis the follwing packages are required:
```{r echo=FALSE}
#required packages

library(raster)
library(sf)
library(sp)
library(rgdal)
library(rgeos)
library(dbscan)
library(dplyr)
library(ggplot2)
library(mapview)

```


## Input data 

Three data set are used:
*HRSL
*OSM
*Admin boundaries

```{r}

#read in spatial data

datapath<-("\\\\srv02\\RL-Institut\\04_Projekte\\240_NESP2\\03-Projektinhalte\\02_Data\\02_Data_from_online_sources\\05_Population_data\\hrsl_nga_v1")
builtup<-raster(file.path(datapath,"hrsl_nga_settlement.tif"))
adm1<-st_read("\\\\srv02\\RL-Institut\\04_Projekte\\240_NESP2\\03-Projektinhalte\\02_Data\\02_Data_from_online_sources\\04_Administrative_Boundaries\\nga_admbnda_adm1_osgof")

adm1 %>% select(admin1Name) %>% ggplot() + geom_sf() + labs(title = "Nigeria")

adm1 %>% filter(admin1Name == "Niger") %>% ggplot() + geom_sf() + labs(title = "Niger")

state_list<-list()


```


## Data processing

Extract the population raster data for each state

```{r echo=FALSE}
Sokoto<- adm1[adm1$admin1Name=='Sokoto', ]

builtup_Sokoto<-crop(builtup,Sokoto) #crop clips the data to the given extent
builtup_Sokoto<-mask(builtup_Sokoto,Sokoto) #mask excludes data inside the extent but outside the specified shape

#transform raster centerpoints to points
pts_buildup_Sokoto<-rasterToPoints(builtup_Sokoto, spatial=TRUE)

summary(pts_buildup_Sokoto)

input=(as.data.frame.array(pts_buildup_Sokoto@coords))


```

## Clustering

The DBSCAN algorithm is used to cluster the points based on the distance to each other and a minimum number of points per cluster.

```{r}

test=dbscan(input, eps = 0.00083, minPts = 3, weights = NULL)
input$cluster <- test$cluster

#split the original data into two according to whether dbscan has assigned or cluster or noise.

groups  <- input %>% filter(cluster != 0)
noise  <- input %>% filter(cluster== 0)

out <- input %>% 
  st_as_sf(coords = c("x","y"), crs = 4326) %>%
  group_by(cluster) %>% 
  summarise() %>% 
  st_convex_hull()

out %>% filter(cluster == 0) %>% plot()

outfiltered <- out %>% filter(cluster != 0)
union <- st_union(outfiltered)



```

## Visualization

The resulting clusters are shown in the map below.


```{r}

mapview(union[[1]], color= "red", col.regions = "red")



```
