---
title: "cluster_analytics"

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
  encoding=encoding,
  output_file=file.path(dirname(input_file), out_dir, 'index.html'))})

author: "Catherina Cader"
date: "11 4 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Overview 

This document provides an overview of the population cluster derived from the High Resolution Layer (<https://www.ciesin.columbia.edu/data/hrsl/>) and OSM (<https://www.openstreetmap.org/data>) for the case of Nigeria.
The data processing and analysis is done using R. This document provides a transparent documentation on the processes and results.

For the analysis the follwing packages are required:
```{recho=FALSE}
#required packages

library(raster)
library(sf)
library(sp)
library(rgdal)
library(rgeos)
library(dbscan)
library(dplyr)

```


## Input data 

Three data set are used:
*HRSL
*OSM
*Admin boundaries

```{r}

#read in spatial data

adm<- st_read("docs/data/nga_admbnda_adm1_osgof")

```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
